Pitfall of Prediction-orienterd Machine Learning Models and Economic Value of Statistical Efficiency in On-farm Precision Experimentation

### 2022-02-20 Writing/manuscript.rmd
result 1st draft

The performances of machine learning and traditional regression models from 1,000 simulations were compared in Figure \@ref(fig:comb-boxplot-all). Three performance measurements: (1) out-of-sample yield prediction, (2) out-of-sample EONR prediction, and (3) profit resultant from the predicted EONR.

Table of winner counts.

Look at the large field (37.3 ha) scenario in panel (A) of Figure \@ref(fig:comb-boxplot-all).

- Indeed, the simulation results showed that machine learning models have better yield predictions than traditional regression models. On average, the out-of-sample predicted yields' RMSE of RF (`r mean_data[field_size=="37.3 ha"&model=="RF", rmse_yield] %>% round(2)` kg ha$^{-1}$) and BRF (`r mean_data[field_size=="37.3 ha"&model=="BRF", rmse_yield] %>% round(2)` kg ha$^{-1}$) models were significantly smaller than that of OLS (`r mean_data[field_size=="37.3 ha"&model=="OLS", rmse_yield] %>% round(2)` kg ha$^{-1}$) or SER (`r mean_data[field_size=="37.3 ha"&model=="SER", rmse_yield] %>% round(2)` kg ha$^{-1}$) model. Those results are consistent with existing literature (**citations**).
- However, the better yield prediction of machine learning models does not translate into better EONR prediction. The average RMSE of out-of-sample predicted EONR of RF (`r mean_data[field_size=="37.3 ha"&model=="RF", rmse_eonr] %>% round(2)` kg ha$^{-1}$) was much larger than that of OLS (`r mean_data[field_size=="37.3 ha"&model=="OLS", rmse_eonr] %>% round(2)` kg ha$^{-1}$) or SER (`r mean_data[field_size=="37.3 ha"&model=="SER", rmse_eonr] %>% round(2)` kg ha$^{-1}$) model. BRF model's EONR prediction (RMSE `r mean_data[field_size=="37.3 ha"&model=="BRF", rmse_eonr] %>% round(2)` kg ha$^{-1}$) was better than RF model, which was similar to OLS model, but was still significantly worse than SER model. Those results showed that, despite machine learning models have better yield prediction, their EONR prediction is worse than traditional regression models.
- Ultimately, the worse EONR prediction of machine learning models resulted in lower production profit compared to the traditional regression models. The 1,000-simulation average profit of RF ($\$$`r mean_data[field_size=="37.3 ha"&model=="RF", profit] %>% round(2)` ha$^{-1}$) or BRF ($\$$`r mean_data[field_size=="37.3 ha"&model=="BRF", profit] %>% round(2)` ha$^{-1}$) model is significantly lower than that of OLS ($\$$`r mean_data[field_size=="37.3 ha"&model=="OLS", profit] %>% round(2)` ha$^{-1}$) or SER ($\$$`r mean_data[field_size=="37.3 ha"&model=="SER", profit] %>% round(2)` ha$^{-1}$) model.

@@

- Within the machine learning models, the RF model performs especially poor in EONR prediction. Its EONR prediction RMSE was much larger than BRF model, and its associated profit was extremely low. 
- Even if we "cheat" and feed RF model with the best possible data (the true yield response parameters), its EONR prediction (RMSE `r mean_data[field_size=="37.3 ha"&model=="RF_b", rmse_eonr] %>% round(2)` kg ha$^{-1}$) and profit performance ($\$$`r mean_data[field_size=="37.3 ha"&model=="RF_b", profit] %>% round(2)` ha$^{-1}$) were still slightly worse than BRF model ($\$$`r mean_data[field_size=="37.3 ha"&model=="BRF", profit] %>% round(2)` ha$^{-1}$) with observed data. 
- Note that RF is currently a very popular machine learning method used by many agricultural studies. Our simulation results showed that although RF may be OK for yield prediction, it is terrible in EONR prediction, and will cause big economic loss if using it in decision making. Even after spending great efforts (and costs) to obtain the perfect information, using RF model still won't achieve good economic performance.

@@

Figure \@ref(fig:comb-boxplot-all) panels (B) and (C) showed the simulation results for medium (18.7 ha) and small (9.3 ha) fields.

- The result pattern of model performances was generally robust to field sizes. For any sized field, the machine learning models still have better yield predictions while worse economic performances than traditional regression models. 
- But we do notice that machine learning models' economic performances catch up when field size decreases. Especially, BRF model outperformed OLS model in EONR prediction and profit performance in small and medium fields. Although BRF model still cannot win SER in profit performance, its profit gap between SER reduced from $\$$`r (mean_data[field_size=="37.3 ha"&model=="SER", profit] - mean_data[field_size=="37.3 ha"&model=="BRF", profit]) %>% round(2)` ha$^{-1}$ in large field to $\$$`r (mean_data[field_size=="18.7 ha"&model=="SER", profit] - mean_data[field_size=="18.7 ha"&model=="BRF", profit]) %>% round(2)` ha$^{-1}$ in medium field and $\$$`r (mean_data[field_size=="9.3 ha"&model=="SER", profit] - mean_data[field_size=="9.3 ha"&model=="BRF", profit]) %>% round(2)` ha$^{-1}$ in small field.
- The results may suggest machine learning models (especially BRF) are more desirable in small fields' EONR predictions. But under the context of current farming practice, most production fields are larger than the smallest field (9.3 ha) scenario of our simulation, SER model is economically best option.

@@

- One simulation example (cherry picking, similar to the mean pattern)

- One location's yield response curves

- Treatment effects

short summary:

## OLS vs. SER: Statistical efficiency is money.

We should be more actively use SER model in yield response estimation.

## EONR prediction and profit is not perfectly one-to-one corresponding. (perhaps too much; remove this point)

Lots of agronomy studies use predicted EONR vs "true" EONR as the criteria to evaluate input recommendation algorithms. ("true" EONR is from yield responses estimated from trial data)

illustrative figure:

@@

```{r comb-boxplot-all, fig.cap="The estimation performances of machine learning models (RF, RF_b, BRF) and traditional regression models (OLS, SER) over 1,000 simulations for: (A) large field (37.3 ha), (B) medium field (18.7 ha), and (C) small field (9.3 ha). The performance measurements include: (1) out-of-sample yield prediction RMSE, (2) out-of-sample EONR prediction RMSE, and (3) profit resultant from the predicted EONR.", echo=FALSE, fig.height=8.5, fig.width=7, cache=FALSE}
(comb_boxplot_large + ggtitle("(A) large field")) / (comb_boxplot_medium + ggtitle("(B) medium field")) / (comb_boxplot_small + ggtitle("(C) small field"))
```

```{r single-simu-yield, fig.cap="Example simulation, yield.", fig.width=6, fig.height=7.5, echo=FALSE, cache=FALSE}
single_simu_yield
```

```{r single-simu-EONR, fig.cap="Example simulation, EONR", fig.width=6, fig.height=7.5, echo=FALSE, cache=FALSE}
single_simu_EONR
```

>>>>

The performances of machine learning and traditional regression models from 1,000 simulations were compared in Figure \@ref(fig:comb-boxplot-all). Three performance measurements were presented: (1) out-of-sample yield prediction, (2) out-of-sample EONR prediction, and (3) profit resultant from the predicted EONR.

First look at the large field (37.3 ha) scenario as shown in panel (A) of Figure \@ref(fig:comb-boxplot-all).

- Indeed, machine learning models showed better yield prediction performances than traditional regression models. On average, the out-of-sample predicted yields' RMSE of RF (`r mean_data[field_size=="37.3 ha"&model=="RF", rmse_yield] %>% round(2)` kg ha$^{-1}$) and BRF (`r mean_data[field_size=="37.3 ha"&model=="BRF", rmse_yield] %>% round(2)` kg ha$^{-1}$) models were significantly smaller than that of OLS (`r mean_data[field_size=="37.3 ha"&model=="OLS", rmse_yield] %>% round(2)` kg ha$^{-1}$) or SER (`r mean_data[field_size=="37.3 ha"&model=="SER", rmse_yield] %>% round(2)` kg ha$^{-1}$) model. Those results are consistent with existing literature (**citations**).
- However, the better yield predictions of machine learning models did not translate into better EONR predictions. The average RMSE of out-of-sample predicted EONR of RF model (`r mean_data[field_size=="37.3 ha"&model=="RF", rmse_eonr] %>% round(2)` kg ha$^{-1}$) was much larger than that of OLS (`r mean_data[field_size=="37.3 ha"&model=="OLS", rmse_eonr] %>% round(2)` kg ha$^{-1}$) or SER (`r mean_data[field_size=="37.3 ha"&model=="SER", rmse_eonr] %>% round(2)` kg ha$^{-1}$) model. BRF model's EONR prediction (RMSE `r mean_data[field_size=="37.3 ha"&model=="BRF", rmse_eonr] %>% round(2)` kg ha$^{-1}$) was better than RF model, and was similar to OLS model, but still significantly worse than SER model. Those simulation results demonstrate that despite machine learning models have better yield predictions, their EONR predictions are worse than traditional regression models.
- Ultimately, the worse EONR prediction of machine learning models resulted in lower production profit compared to the traditional regression models. The 1,000-simulation average profit of RF ($\$$`r mean_data[field_size=="37.3 ha"&model=="RF", profit] %>% round(2)` ha$^{-1}$) or BRF ($\$$`r mean_data[field_size=="37.3 ha"&model=="BRF", profit] %>% round(2)` ha$^{-1}$) model was significantly lower than that of OLS ($\$$`r mean_data[field_size=="37.3 ha"&model=="OLS", profit] %>% round(2)` ha$^{-1}$) or SER ($\$$`r mean_data[field_size=="37.3 ha"&model=="SER", profit] %>% round(2)` ha$^{-1}$) model.

@@

- Within the machine learning models, the RF model performed especially poorly in EONR prediction. Its EONR prediction RMSE was much larger than BRF model, and its associated profit was extremely low (as shown in Figure \@ref(fig:comb-boxplot-all) panel (A)). 
- Even if we "cheat" and feed RF model with the best possible data (i.e., the true yield response parameters), its EONR profit ($\$$`r mean_data[field_size=="37.3 ha"&model=="RF_b", profit] %>% round(2)` ha$^{-1}$) was still slightly lower than BRF model ($\$$`r mean_data[field_size=="37.3 ha"&model=="BRF", profit] %>% round(2)` ha$^{-1}$) that was estimated using observed messy data. 
- It is noteworthy to mention that RF model is currently a very popular machine learning method used in many agricultural studies. Our simulation results showed that although RF may be OK for yield prediction, it is terrible in EONR prediction, and will cause big economic loss if using it in farming recommendation development. Even after spending great efforts (and financial costs) to obtain the perfect information, using RF model still won't achieve good economic performance.

@@

How robust are the results to different field sizes? Figure \@ref(fig:comb-boxplot-all) panels (B) and (C) showed the simulation results for medium (18.7 ha) and small (9.3 ha) sized fields.

- The result pattern that machine learning models have better yield predictions but worse economic performances compared to traditional regression models was generally robust to field sizes. 
- But machine learning models' economic performances catch up when field size decreases. Especially, BRF model outperformed OLS model in EONR prediction and profit performance in small and medium fields. Though BRF model's profit was still lower than SER, the profit gap between BRF and SER reduced from $\$$`r (mean_data[field_size=="37.3 ha"&model=="SER", profit] - mean_data[field_size=="37.3 ha"&model=="BRF", profit]) %>% round(2)` ha$^{-1}$ in large field to $\$$`r (mean_data[field_size=="18.7 ha"&model=="SER", profit] - mean_data[field_size=="18.7 ha"&model=="BRF", profit]) %>% round(2)` ha$^{-1}$ in medium field and $\$$`r (mean_data[field_size=="9.3 ha"&model=="SER", profit] - mean_data[field_size=="9.3 ha"&model=="BRF", profit]) %>% round(2)` ha$^{-1}$ in small field.
- The results may suggest machine learning models (especially BRF) apply better in small fields' data analysis. But since under the context of current farming practice most production fields are larger than the smallest field (9.3 ha) scenario in our simulation, SER model appears to be the economically best option.

@@

Why good yield prediction does not necessarily end up with good economic performance? That sound counterintuitive. We attempt to provide some explanations. Note this is not a statistical paper and providing the rigorous statistical proof is beyond the scope of this paper. We just try providing some illustrative explanations.

- Figure \@ref(fig:single-simu-illustrate) showed the predicted versus true yield (panel A) and EONR (panel B) of the 1,440 locations ($aunit$) from one example simulation's results (large field). The RMSE measurements associated with each model's results for this specific simulation were displayed in the figure as well. Though the machine learning models (RF, RF_b, and BRF) generated well-predicted yields such that the plotted points were well distributed along the 45 degree line, their predicted EONR were not continuous but fell into several discrete values. That results were to a large extent caused by the discrete data of five nitrogen trial rates. But that is how the on-farm experimental data generally look like in practice. On the other hand, the SER model's predicted EONR were much better aligned the 45 degree line, despite its predicted yields were not as good as the machine learning models. The OLS model's EONR predictions are also continuous, but with too big dispersion. 
- We further look into the yield response curve estimation at individual locations. Figure \@ref(fig:yield-curves) illustrated the estimated yield response curves by different models at a specific location ($aunit$). The circle point denoted the trial N rate and true (deterministic part of) yield. This figure showed that the machine learning models' predicted yields were closer to the true yield. However, the shapes of yield response curves predicted by machine learning models were much less accurate. On the other hand, SER model's predicted yield levels were heavily biased, but its predicted yield response curve had the shape much more resembling the true response curve. The EONR is determined by the yield response (mathematically, the slope of the curve), rather than the yield level. 
- Mathematically the slope of the curve: economics term, marginal effect, but the derivative measure is not commonly used in agronomy studies; Agronomy term, treatment effect, which is not the continuous derivative measure, but regarding the different N levels as discrete values. A treatment effect is defined as the change in yield from one N rate to the next level N rate. Here we have five N treatment rates, and therefore define four treatment effects. The estimated versus true treatment effects of N rates for the same simulation round's 1,440 locations were shown in Figure \@ref(fig:treatment-effect) for different models. Notice that we used the quadratic-plateau function as the true yield response in the simulation, and therefore at higher N levels the yields are more likely to approach or pass the plateau and the true treatment effect is quite small or even zero. As can be seen, the machine learning models mostly underestimated the treatment effects at all N levels. That is the major reason why machine learning models failed to predict accurate EONR. In contrast, SER model did a much better job in treatment effect predictions, particularly at the low N levels. OLS model's estimated treatment effects were actually less biased compared to machine learning models, but the variations were just very large.
- (Maybe we don't need to say this, as this is so obvious.) Caveat: The single simulation round we choose was a little cherry-picking. We chose that the round where the five models' profit results were roughly corresponding to the means as shown in Figure \@ref(fig:comb-boxplot-all). Indeed, there are lots of variations across simulation rounds, such that in come simulations machine learning models outperformed SER model. We believe the one example simulation we showed here is representative, and can make a good illustration of the yield versus EONR prediction performances.

## OLS vs. SER: Statistical efficiency is money

(I feel this part is only a minor point in this study. Perhaps it's not important enough to be put in the title.)

- It is also interesting to compare the OLS and SER results. The good performance of SER may imply some good suggestion to production.
- It has been well known in the econometric literature that SER model can improve the estimation efficiency compared to OLS at the presence of unknown spatial data autocorrrelation, which is a very common in crop field characteristics data. 
- However, since it is only an efficiency issue, and SER does not correct for bias, this advantage of SER model has not received enough attention. The efficiency improvement is only regarded as a statistical property, while its economic benefit is not well recognized.
- Our simulation evidence showed that the statistical efficiency of SER model can actually translate into considerable more profitable EONR recommendations. The profit of SER model is higher than OLS model by $\$$`r (mean_data[field_size=="37.3 ha"&model=="SER", profit] - mean_data[field_size=="37.3 ha"&model=="OLS", profit]) %>% round(2)` ha$^{-1}$, $\$$`r (mean_data[field_size=="18.7 ha"&model=="SER", profit] - mean_data[field_size=="18.7 ha"&model=="OLS", profit]) %>% round(2)` ha$^{-1}$, and $\$$`r (mean_data[field_size=="9.3 ha"&model=="SER", profit] - mean_data[field_size=="9.3 ha"&model=="OLS", profit]) %>% round(2)` ha$^{-1}$ for large, medium, and small fields, respectively.
- So, using SER model to update the OLS model is likely to be economically beneficial in on-farm experimental data analysis, and should be encouraged.

## Robustness to yield generating process

- The true yield generating process used in our simulation, quadratic-plateau function, is the most established corn yield response pattern demonstrated in the agronomy literature. But there are also some other corn response functional forms found in existing studies. In case somebody criticize us on this point, we also tried several other corn yield response functions: Mitscherlich‐Baul, quadratic, ...

@@

```{r comb-boxplot-all, fig.cap="The estimation performances of machine learning models (RF, RF_b, BRF) and traditional regression models (OLS, SER) over 1,000 simulations for: (A) large field (37.3 ha), (B) medium field (18.7 ha), and (C) small field (9.3 ha). The performance measurements include: (1) out-of-sample yield prediction RMSE, (2) out-of-sample EONR prediction RMSE, and (3) profit resultant from the predicted EONR.", fig.height=8.5, fig.width=7, echo=FALSE, cache=FALSE}
(comb_boxplot_large + ggtitle("(A) large field")) / (comb_boxplot_medium + ggtitle("(B) medium field")) / (comb_boxplot_small + ggtitle("(C) small field"))
```

```{r single-simu-illustrate, fig.cap="The estimation performances of one example simulation, large field (37.3 ha). Each dot represents a location of the field.", fig.height=8.5, fig.width=7, echo=FALSE, cache=FALSE}
(single_simu_yield + ggtitle("(A)")) / (single_simu_EONR + ggtitle("(B)"))
```

```{r yield-curves, fig.cap="Estimated yield response curves at a specific location", fig.height=8.5, fig.width=7, echo=FALSE, cache=FALSE}
include_graphics(here("GitControlled", "Graphs", "802_yield.png"))
```

```{r treatment-effect, fig.cap="", echo=FALSE, fig.height=8.5, fig.width=7, cache=FALSE}
include_graphics(here("GitControlled", "Graphs", "treatment_sim_58.png"))
```
